---
title: "Assignment 1"
author: 'Enric Reverter & Gerard Pons'
date: "14/10/2021"
output:
  word_document: default
  html_document:
    df_print: paged
  pdf_document: default
---

```{r setup, include=FALSE, eval=F}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=F}
## Data manipulation
require(tidyverse)
require(dplyr)
require(mice)
require(Hmisc)

## Statistics
library(lsr)
library(missMDA)
library(VIM)
library(chemometrics)
library(arules)
require(skimr)
library(lmtest)
require(car)
require(MASS)


## Plots
require(ggplot2)
require(ggExtra)
require(ggthemes)
require(processx)
require(plotly)

require(cowplot)
require(gridExtra)

require(RColorBrewer)
library(viridis)

theme_set(theme_bw())

## Set data path
setwd("..")
data_path = file.path(getwd(), "data")
plot_path = file.path(getwd(), "plots")
```

## Get the data

```{r, eval=T}
# Read datasets
df = read.csv(file.path(data_path, "cars.csv"))
```

### Question 1

**Q: Determine if the response variable (price) has an acceptably normal distribution. Address test to discard serial correlation.**

```{r}
p1 = ggplot(data = df, aes(x=price)) +
  geom_histogram(aes(y=..density..), color = 1, fill = "white", bins = 50, ) +
  geom_density(color = "red", lwd = 1) +
  labs(title = "Histogram & Density for Price")

p2 = ggplot(data = df, aes(sample=price)) +
  geom_qq(shape = 1, color = "royalblue") + geom_qq_line() +
  labs(title = "QQ Plot for Price")

plot_grid(plotlist = list(p1,p2), align = "h", ncol = 2)
```

```{r}
shapiro.test(df$price)
```

```{r}
acf(df$price)
dwtest(price ~ 1, data = df)
```


### Question 2

**Q: Indicate by exploration of the data which are apparently the variables most associated with the response variable (use only the indicated variables).**

```{r}
library(corrplot)
correlation = cor(df[, sapply(df, is.numeric)],method="spearman")
corrplot(correlation, order="AOE", method="color", addCoef.col = "gray10", tl.col = "black", tl.cex = .8, number.digits = 2, number.cex = 0.7, title = "Spearman Correlation Plot", mar=c(0,0,1.5,0))

cor.test(df$price, df$tax,method="spearman")
cor.test(df$price, df$mpg,method="spearman")
cor.test(df$price, df$age,method="spearman")
cor.test(df$price, df$mileage,method="spearman")
```

```{r}
kruskal.test(df$price,df$manufacturer)
kruskal.test(df$price, df$transmission)
kruskal.test(df$price, df$fuelType)
kruskal.test(df$price, df$engineSize_int)
```

### Question 3

**Q: Define a polytomic factor f.age for the covariate car age according to its quartiles and argue if the average price depends on the level of age. Statistically justify the answer.**

```{r}
df$f.age = discretize(df$age,method = "frequency",breaks=4)
table(df$f.age)
round(prop.table(table(df$f.age))*100)

tapply(df$price, df$f.age, mean)
Boxplot(df$price~df$f.age)

is_outlier = function(x) {
  return(x < quantile(x, 0.25) - 3 * IQR(x) | x > quantile(x, 0.75) + 3 * IQR(x))
}

df %>%
  group_by(f.age) %>%
  tibble::rownames_to_column(var = "id") %>%
  mutate(outlier = ifelse(is_outlier(price), id, NA)) %>%
  ggplot(., aes(x = f.age, y = price, fill = f.age)) +
    geom_boxplot(alpha = 0.6, outlier.alpha = NULL) +
    scale_fill_brewer(palette = "Blues") +
    geom_text(aes(label = outlier), na.rm = TRUE, hjust = -0.3) +
    labs(title = "Boxplots of price by f.age",
         caption = "Only severe outliers are labeled") +
    theme(legend.position = "none")
```

```{r}
kruskal.test(df$price~df$f.age)

pairwise.wilcox.test(df$price,df$f.age,alternative='less')
```

### Question 4

**Q: Calculate and interpret the anova model that explains car price according to the age factor and the fuel type.**

```{r}
options(contrasts=c('contr.treatment','contr.treatment'))

m0 = lm(df$price~1)
m1 = lm(df$price~df$f.age)
m2 = lm(df$price~df$fuelType)
m3 = lm(df$price~df$f.age+df$fuelType)
m4 = lm(df$price~df$f.age*df$fuelType)

anova(m0,m3) #General effect of the proposed model
anova(m3) 
summary(m3)
```

### Question 5

**Q: Do you think that the variability of the price depends on both factors? Does the relation between price and age factor depend on fuel type?**

```{r}
anova(m0,m1) #Gross effect factor age
anova(m0,m2) #Gross effect fuelType

anova(m3,m2) #Net fuelType effect
anova(m3,m1) #Net factor age effect

anova(m3,m4) #Check interactions
```

```{r}
par(mfrow=c(1,1))
interaction.plot(df$f.age,df$fuelType,df$price)
```

### Question 6 & 7

**Q: Calculate the linear regression model that explains the price from the age: interpret the regression line and assess its quality. What is the percentage of the price variability that is explained by the age of the car?**

```{r}
m2 = lm(df$price~df$age)
```

```{r}
ggplot(df, aes(x = age, y = price)) +
  geom_point(alpha = 0.8) +
  stat_smooth(method = "lm", col = "royalblue") +
  labs(title = "Linear Model: price ~ age")
```

```{r}
summary(m2)
par(mfrow=c(2,2))
plot(m2)
par(mfrow=c(1,1))

boxcox(df$price ~ df$age)
```

### Question 8

**Q: Do you think it is necessary to introduce a quadratic term in the equation that relates the price to its age?**

```{r}
m22 = lm(df$price~df$age+I(df$age^2))
summary(m22)
anova(m2,m22)

par(mfrow=c(2,2))
plot(m22)
par(mfrow=c(1,1))
```

```{r}
boxTidwell(df$price~I(df$age+0.1))
```

### Question 9

**Q: Are there any additional explanatory numeric variables needed to the car price? Study collinearity effects.**

```{r}
m3 = lm(df$price~df$age+df$mileage+df$tax+df$mpg)
summary(m3)
Anova(m3)
vif(m3)
```

```{r}
m3 = lm(df$price~df$age+df$mpg)
summary(m3)
vif(m3)
boxTidwell(df$price~I(df$age+0.1)+df$mpg)
```

```{r}
m33 = lm(df$price~ df$age + I(df$mpg^-2))
summary(m33)
vif(m33)
```

```{r}
BIC(m3, m33)
```

```{r}
# scatter3d(df$price ~ df$age + I(df$mpg^-2), groups = as.factor(df$fuelType), grid = F)
```

### Question 10

**Q: After controlling by numerical variables, indicate whether the additive effect of the available factors on the price are statistically significant.**

```{r}
m12= lm(df$price~df$age+I(df$mpg^-2)+df$manufacturer+df$engineSize_int+df$transmission+df$fuelType)
full_model = step(m12,k=log(nrow(df)))
summary(full_model)
anova(m3,full_model)
```

```{r}
avPlots(m12)
```

  ### Question 11

**Q: Select the best model available so far. Interpret the equations that relate the explanatory variables to the answer (rate).**

```{r}
summary(full_model)
```

```{r}
par(mfrow=c(2,2))
plot(full_model)
par(mfrow=c(1,1))
```

### Question 12

**Q: Study the model that relates the logarithm of the price to the numerical variables.**

```{r}
mlog = lm(log(df$price)~df$age+df$mileage+df$tax+df$mpg)
summary(mlog)
Anova(mlog)
vif(mlog)
```

```{r}
mlog = lm(log(df$price)~df$age+df$tax+df$mpg)
summary(mlog)
```

```{r}
boxTidwell(log(df$price)~I(df$age+0.1)+I(df$tax+0.1)+df$mpg)

mlog1 = lm(log(df$price)~df$age+df$tax+I(df$mpg^-2))

anova(mlog,mlog1)
BIC(mlog, mlog1)

boxTidwell(log(df$price)~I(df$age+0.1)+I(df$tax+0.1)+I(df$mpg^-2))

mlog2 = lm(log(df$price)~I(df$age^2)+df$tax+I(df$mpg^-2))

anova(mlog1,mlog2)
BIC(mlog1, mlog2)
```

```{r}
par(mfrow=c(2,2))
plot(mlog1)
par(mfrow=c(1,1))
```


### Question 13

**Q: Once explanatory numerical variables are included in the model, are there any main effects from factors needed?**

```{r}
mlog = lm(log(df$price)~df$age+df$tax+I(df$mpg^-2)+df$engineSize_int+df$manufacturer+df$transmission+df$fuelType)
logmodel = step(mlog, k = log(nrow(df)))
summary(logmodel)
vif(logmodel)
```

### Question 14

**Q: Graphically assess the best model obtained so far.**

```{r}
par(mfrow=c(2,2))
plot(logmodel) 
par(mfrow=c(1,1))
```

```{r}
avPlots(logmodel)
marginalModelPlots(logmodel, terms = ~df$age+df$tax+I(df$mpg^-2))
```

### Question 15

**Q: Assess the presence of outliers in the studentized residuals at a 99% confidence level. Indicate what those observations are.**

```{r}
# mb is best model
mb = logmodel
n = dim(df)[1]
p = mb$rank
res_mb = rstudent(mb)
cut_off = qt(0.995,n-p-1)

Boxplot(res_mb)
abline(h=cut_off,col=2)
abline(h=-cut_off,col=2)

nrow(df[which(abs(res_mb)>cut_off),])

aux = df[which(abs(res_mb)>cut_off),]
summary(aux)

table(aux$model)
prop.table(table(aux$transmission)); prop.table(table(df$transmission))
prop.table(table(aux$fuelType)); prop.table(table(df$fuelType))
prop.table(table(aux$engineSize_int)); prop.table(table(df$engineSize_int))
prop.table(table(aux$manufacturer)); prop.table(table(df$manufacturer))
mean(aux$price); mean(df$price)
mean(aux$age); mean(df$age)
mean(aux$tax); mean(df$tax)
mean(aux$mpg); mean(df$mpg)
```

### Question 16

**Q: Study the presence of a priori influential data observations, indicating their number according to the criteria studied in class.**

```{r}
hat = hatvalues(mb)
hat_cut = 3*p/n #if we assume the dataset is small, then 2p/n

Boxplot(hat)
abline(h=hat_cut,col=2)

sum(hat>hat_cut)

aux = df[which(hat>hat_cut),]

table(aux$model)
prop.table(table(aux$transmission)); prop.table(table(df$transmission))
prop.table(table(aux$fuelType)); prop.table(table(df$fuelType))
prop.table(table(aux$engineSize_int)); prop.table(table(df$engineSize_int))
prop.table(table(aux$manufacturer)); prop.table(table(df$manufacturer))
mean(aux$price); mean(df$price)
mean(aux$age); mean(df$age)
mean(aux$tax); mean(df$tax)
mean(aux$mpg); mean(df$mpg)
```

### Question 17

**Q: Study the presence of a posteriori influential values, indicating the criteria studied in class and the actual atypical observations.**

```{r}
cook = cooks.distance(mb)
cook_cut = 4/(n-p)

Boxplot(cook)
abline(h=cook_cut)

nrow(df[which(cook>cook_cut),])

aux = df[which(cook>cook_cut),]

table(aux$model)
prop.table(table(aux$transmission)); prop.table(table(df$transmission))
prop.table(table(aux$fuelType)); prop.table(table(df$fuelType))
prop.table(table(aux$engineSize_int)); prop.table(table(df$engineSize_int))
prop.table(table(aux$manufacturer)); prop.table(table(df$manufacturer))
mean(aux$price); mean(df$price)
mean(aux$age); mean(df$age)
mean(aux$tax); mean(df$tax)
mean(aux$mpg); mean(df$mpg)
```

```{r}
df[which(cook>0.5),]
```

```{r}
influencePlot(mb)
```

### Question 18

**Q: Given a 5-year old car, the rest of numerical variables on the mean and factors on the reference level, what would be the expected price with a 95% confidence interval?**

```{r}
mpred = lm(log(price)~age+mpg+engineSize_int+manufacturer+transmission+fuelType,data=df)
pred_df<-data.frame(age=5,tax=mean(df$tax),mpg=mean(df$mpg),engineSize_int='[0.0,1.5)',manufacturer='Audi',
                    transmission='Automatic',fuelType='Diesel')
exp(predict(mpred,pred_df,conf=0.95,interval=c("pred"),se.fit=T)$fit)
```

### Question 19

**Q: Summarize what you have learned by working with this interesting real dataset.**

### Plots

```{r}
ggplot(data = df, aes(x="", y=price)) +
  geom_boxplot(width=0.5) +
  scale_y_continuous(labels=scales::comma) +
  labs(title="Wrap of Price by Manufacturer",
       y="Price [€]") +
  # Do not show x axis
  theme(axis.text.x=element_blank(), axis.ticks.x = element_blank(), axis.line.x = element_blank(), axis.title.x=element_blank()) +
  facet_wrap(vars(manufacturer))

ggplot(data = df, aes(x="", y=age)) +
  geom_boxplot(width=0.5) +
  scale_y_continuous(labels=scales::comma) +
  labs(title="Wrap of Age by Manufacturer",
       y="Age [Years]") +
  # Do not show x axis
  theme(axis.text.x=element_blank(), axis.ticks.x = element_blank(), axis.line.x = element_blank(), axis.title.x=element_blank()) +
  facet_wrap(vars(manufacturer))

ggplot(data = df, aes(x="", y=price)) +
  geom_boxplot(width=0.5) +
  scale_y_continuous(labels=scales::comma) +
  labs(title="Wrap of Price by Engine Size",
       y="Price [€]") +
  # Do not show x axis
  theme(axis.text.x=element_blank(), axis.ticks.x = element_blank(), axis.line.x = element_blank(), axis.title.x=element_blank()) +
  facet_wrap(vars(engineSize_int))

ggplot(data = df, aes(x="", y=price)) +
  geom_boxplot(width=0.5) +
  scale_y_continuous(labels=scales::comma) +
  labs(title="Wrap of Price by Fuel Type",
       y="Price [€]") +
  # Do not show x axis
  theme(axis.text.x=element_blank(), axis.ticks.x = element_blank(), axis.line.x = element_blank(), axis.title.x=element_blank()) +
  facet_wrap(vars(fuelType))

ggplot(data = df, aes(x="", y=price)) +
  geom_boxplot(width=0.5) +
  scale_y_continuous(labels=scales::comma) +
  labs(title="Wrap of Price by Transmission",
       y="Price [€]") +
  # Do not show x axis
  theme(axis.text.x=element_blank(), axis.ticks.x = element_blank(), axis.line.x = element_blank(), axis.title.x=element_blank()) +
  facet_wrap(vars(transmission))
```

```{r}
# pairwise.wilcox.test(df$age, df$manufacturer, alternative = "less")
# pairwise.wilcox.test(df$mileage, df$manufacturer, alternative = "less")
# pairwise.wilcox.test(df$mileage, df$manufacturer, alternative = "less")
# table(df$manufacturer, df$transmission)
# table(df$manufacturer, df$engineSize_int)
# table(df$manufacturer, df$fuelType)
```

